---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
---
The future border crossings from Ukraine to Slovakia are predicted and estimated using a K-NN regression algorithm. The dataset is taken from the daily updated UNHCR border crossing data, which can be found at https://data.unhcr.org/en/situations/ukraine/location/10785 . The forecast is contingent on the relative stability of the time series, and hence may not be able to take into account uncertainties that are implicit in mass social phenomena. 

```{r}
install.packages(c("tidymodels","modeltime","dplyr","timetk", "ggplot2", "glmnet", "tidyquant", "magrittr", "kknn"))
```


```{r}
library(tidymodels)
library(dplyr)
library(timetk)
library(magrittr)
library(workflows)
library(parsnip)
library(recipes)
library(yardstick)
library(glmnet)
library(tidyverse)
library(tidyquant)
library(timetk)
library(kknn)
```


```{r}
BC <- read.csv("BorderCrossingsFromUKR.csv") %>% select(timeseries.data_date, timeseries.individuals) %>% set_names(c("date","people"))
BC$date = as.Date(BC$date, format = '%Y-%m-%d')
BC
```


```{r}
BC %>%
    ggplot(aes(x = date, y = people)) +
    geom_rect(xmin = as.numeric(ymd("2022-02-01")),
              xmax = as.numeric(ymd("2023-04-01")),
              ymin = 0, ymax = 10000,
              fill = palette_light()[[4]], alpha = 0.01) +
    geom_point(alpha = 0.5, color = palette_light()[[1]]) 
```


```{r}
train_BC <- BC %>% filter(date < ymd("2023-01-01") & date >= ymd("2022-04-01"))
test_BC <- BC %>% filter(date >= ymd("2023-01-01"))

train_BC
```


```{r}
# Add time series signature
recipe_spec_timeseries <- recipe(people ~ date, data = train_BC) %>%
    step_timeseries_signature(date) %>%   #adding time series signatures
    step_rm(date) %>%
    step_rm(contains("iso"), 
        contains("minute"), 
        contains("hour"),
        contains("am.pm"), 
        contains("xts")) %>%
    step_zv(date_year)%>%
    step_dummy(contains("lbl"), one_hot = TRUE) 
```


```{r}
bake(prep(recipe_spec_timeseries), new_data = train_BC)
```


```{r}
model_spec_glmnet <-  nearest_neighbor(neighbors = 3, weight_func = "gaussian") %>%
  # Using K-NN Regression
  set_mode("regression") %>%
  set_engine("kknn")
```


```{r}
workflow_glmnet <- workflow() %>%
    add_recipe(recipe_spec_timeseries) %>%
    add_model(model_spec_glmnet)
    
workflow_glmnet
```


```{r}
workflow_trained <- workflow_glmnet %>% fit(data = train_BC)
```


```{r}
prediction_BC <- workflow_trained %>% 
    predict(test_BC) %>%
    bind_cols(test_BC) 
    
prediction_BC
```


```{r}
prediction_BC %>% metrics(people, .pred)
```


```{r}
idx <- BC %>% tk_index()
BC_sum <- idx %>% tk_get_timeseries_summary()
```


```{r}
idx_future <- idx %>% tk_make_future_timeseries(length_out = 365)
future_BC <- tibble(date = idx_future) 

future_BC
```


```{r}
future_predictions_BC <- workflow_glmnet %>% 
    fit(data = BC[BC$date >= "2022-04-01",]) %>%
    predict(future_BC) %>%
    bind_cols(future_BC)
```

```{r}
BC %>%
    ggplot(aes(x = date, y = people)) +
    # future data
    geom_point(aes(x = date, y = .pred), data = future_predictions_BC,
               alpha = 0.5, color = palette_light()[[2]]) +
    geom_smooth(aes(x = date, y = .pred), data = future_predictions_BC,
                method = 'loess') + 
    labs(title = "Border Crossing Forecast", x = "") +
    theme_tq()
```


```{r}
test_resid_sd <- prediction_BC %>%
    summarize(stdev = sd(people - .pred))

future_predictions_BC <- future_predictions_BC %>%
    mutate(
        lo.95 = .pred - 1.96 * test_resid_sd$stdev,
        lo.80 = .pred - 1.28 * test_resid_sd$stdev,
        hi.80 = .pred + 1.28 * test_resid_sd$stdev,
        hi.95 = .pred + 1.96 * test_resid_sd$stdev
    )
```


```{r}
options(repr.plot.width=20, repr.plot.height=10)
BC %>%
    ggplot(aes(x = date, y = people)) +
    geom_point(alpha = 0, color = palette_light()[[1]]) +
    geom_line(alpha = 0.9)+
    geom_ribbon(aes(y = .pred, ymin = lo.95, ymax = hi.95 ), 
                data = future_predictions_BC, 
               fill = "#D5DBFF", color = NA, size = 0) +
   # geom_ribbon(aes(y = .pred, ymin = lo.80, ymax = hi.80,fill = key), 
    #            data = future_predictions_BC,
     #          fill = "#596DD5", color = NA, size = 0, alpha = 0.6) +
    geom_point(aes(x = date, y = .pred), data = future_predictions_BC,
               alpha = 0, color = palette_light()[[2]]) +
     geom_line(aes(x = date, y = .pred), data = future_predictions_BC,
             alpha = 0.8, color = "red") +
   # geom_smooth(aes(x = date, y = .pred), data = future_predictions_BC,
    #            method = 'loess', color = "red", alpha=0.1) + 
    labs(title = "Border Crossings Into Slovakia From Ukraine", x = "", caption='Shaded region indicates 95% CI.') +
    theme_tq()
```


```{r}
future_predictions_BC
```


